---
title: "Simulating a fished population"
author: "Jaideep Joshi"
date: "28 March 2022"
output: html_document
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = F, echo = F}
library(rfish)
library(tidyverse)
source("../tests/ref/parameters.cod.R")
source("../tests/ref/simulator.7.R")
```


To simulate a fished population, it is essential to first simulate an unfished population until equilibrium. The properties of such an unfished equilibriated population must be used to simulate a fished population. 

### Using a simulator for fish populations 

The `rfish` package provides a simulator to easily simulate populations of fish without the hassle of manual updating. The simulator can also simulate multiple populations with different control parameters at once, making it very convenient to analyse the control space (we will look at the control space in a later tutorial).  

As usual, we begin with creating a prototype fish 

```{r}
# Create a prototype fish as usual
fish = new(Fish)
```

We then create a simulator, specifying the prototype fish:

```{r}
sim = new(Simulator, fish)
```

The simulator already contains a reference population, which we can simulate under non-fished conditions. The simulator will perform requisite calculations on this population as and when required.

```{r}
sim$equilibriateNaturalPopulation(1.93e3, 5.61)
```


### Simulating a fished population

Finally, we are ready to simulate a fished population. We create a population to be simulated and set its parameters

```{r}
pop = new(Population, fish)
```

Then, suppose we want to simulate this population for 200 years with a harvest proportion of 0.3 and minimum size limit of 45, 

```{r}
nsteps = 200
h = 0.3
lf = 45
```

we simply do
```{r}
pop$verbose = T
# pop$set_superFishSize(1e6*(exp(-6*h)))
pop$set_superFishSize(20e6)
res_ibm = sim$simulate(pop, lf, h, nsteps, 1.93e3, 5.61, T)
```

The last argument is a flag to re-initialize the population before simulating. If this flag is set to false, simulation will continue from the current state of the population, allowing us to simulate different control regimes over time. 

Let's plot the output from the simulation (compared with the model of Dankel et al.XXX)

```{r}
res = simulate(h, lf, F)

d = pop$get_state()
table(d$age)
par(mfrow = c(2,3), mar=c(4,4,1,1))

ssb.max = max(c(res_ibm$ssb/1e9, res$summaries$SSB/1e9))
plot(y=res_ibm$ssb/1e9, x=seq(1,nsteps,1), ylab="SSB (MT)", xlab="Year", col="cyan3", type="l", ylim=c(0,ssb.max))
points(y=res$summaries$SSB/1e9, x=res$summaries$year, type="l")

yield.max = max(c(res_ibm$yield/1e9, res$summaries$Y/1e9))
plot(y=res_ibm$yield/1e9, x=seq(1,nsteps,1), ylab="Yield (MT)", xlab="Year", col="cyan3", type="l", ylim=c(0,yield.max))
points(y=res$summaries$Y/1e9, x=res$summaries$year, type="l")

emp.max = max(c(res_ibm$employment.sea, res$summaries$D.sea))
plot(y=res_ibm$employment.sea, x=seq(1,nsteps,1), ylab="Employment at sea (person-years)", xlab="Year", col="cyan3", type="l", ylim=c(0,emp.max))
points(y=res$summaries$D.sea, x=res$summaries$year, type="l")

emp.max = max(c(res_ibm$employment.shore, res$summaries$D.shr))
plot(y=res_ibm$employment.shore, x=seq(1,nsteps,1), ylab="Employment on shore (person-years)", xlab="Year", col="cyan3", type="l", ylim=c(0,emp.max))
points(y=res$summaries$D.shr, x=res$summaries$year, type="l")

p.sea.max = max(c(res_ibm$profit.sea/1e9, res$summaries$P.sea/1e9))
p.sea.min = min(c(res_ibm$profit.sea/1e9, res$summaries$P.sea/1e9))
plot(y=res_ibm$profit.sea/1e9, x=seq(1,nsteps,1), ylab="Profit at sea (Billions NOK)", xlab="Year", col="cyan3", type="l", ylim=c(p.sea.min,p.sea.max))
points(y=res$summaries$P.sea/1e9, x=res$summaries$year, type="l")

p.shr.max = max(c(res_ibm$profit.shore/1e9, res$summaries$P.shr/1e9))
p.shr.min = min(c(res_ibm$profit.shore/1e9, res$summaries$P.shr/1e9))
plot(y=res_ibm$profit.shore/1e9, x=seq(1,nsteps,1), ylab="Profit at shore (Billions NOK)", xlab="Year", col="cyan3", type="l", ylim=c(p.shr.min,p.shr.max))
points(y=res$summaries$P.shr/1e9, x=res$summaries$year, type="l")

par(mfrow=c(1,1))
d = pop$get_state()
d1 = d %>% group_by(age) %>% summarize(mat = length(which(isMature))/length(isMature))
plot(mat, type="l")
points(d1$mat~d1$age, type="o", col="cyan3")
```