---
title: "Growth model calibration"
author: "Jaideep Joshi"
date: "9 May 2022"
output: html_document
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = F, echo = F}
library(rfish)
```

### Growth model parameters

```{r}
beta = 0.655793;
alpha = 3.056863227;
c = 6.318308;
k = 0.00674;
# r = 0.090367;
# q = 1;

l0 = 9.1

gamma1 = 1-beta;
gamma2 = alpha;
# alpha1 = c;
alpha2 = k;
# gsi = r/q;

beta1 = 2.430;
beta2 = -7.07e-5;
beta3 = 0.178;
Tmean = 5.61;
tsbmean = 1.93e9/1e6;

```


```{r}
dl_power = function(tsb, temp, alpha1){
	t2 = gamma1 * alpha1 * alpha2^(-gamma1)
	t2 = t2 * exp(beta2*(tsb-tsbmean) + beta3*(temp-Tmean))
	t2
}

length_juvenile = function(init_length, tsb, temp, alpha1){
  t1 = init_length^ (gamma1 * gamma2)
  t2 = dl_power(tsb, temp, alpha1)
  l_new = (t1 + t2)^ (1 / (gamma1 * gamma2))
  l_new
}
```


### Simulate growth over five age classes


```{r}
fishLen = function(alpha1, tsb = tsbmean, temp = Tmean, ny=6){
  years = (1:ny)+1
  len = numeric(ny+1)
  len[1] = l0
  for (i in years){
    len[i] = length_juvenile(len[i-1], tsb, temp, alpha1)
  }
  len[-1]
}

fishGrowth = function(alpha1, tsb = tsbmean, temp = Tmean, ny=6){
  diff(fishLen(alpha1, tsb, temp, ny))
}

ny=6

plot(fishLen(c)~seq(1,ny), xlab="Age", ylab="Length", type="l", lwd=2)
```


### Compare with C++ sim to ensure all is well

```{r}
# NEW MODEL
N = 100
plot(x=1, y=NA, xlim=c(0,31), ylim=c(0,200), xlab="age", ylab = "length")
for (f in 1:N){
  fish = new(Fish)
  fish$par$s0 = 0.09637
  fish$par$Bhalf_growth = 1e11

  dat = data.frame(t_birth=0,	age=0,	isMature=0,	isAlive=0,	length=0,	weight=0)
  dat = rbind(dat, fish$get_state())
  dat = dat[-1,]
  for (i in 1:30){
    dat = rbind(dat, fish$get_state())
    fish$updateMaturity()
    fish$grow(0)
    fish$set_age(fish$age+1)
  }
  points(dat$length~dat$age, type="l", col=rainbow(N)[f])
}
points(fishLen(c)~seq(1,ny), xlab="Age", ylab="Length", type="l", lwd=2)
```

### Comparison of mean predicted and observed dl, with default value of alpha1


```{r}
# Statistical model
tsb_ano = 0
temp_ano = 0
dl_obs = exp(beta1+beta2*tsbmean + beta2*tsb_ano + beta3*temp_ano)

plot(diff(fishLen(c))~seq(1,ny-1), xlab="Age", ylab="dL", type="l", lwd=2, ylim=c(9,10))
points((fishGrowth(c))~seq(1,ny-1), col="red", xlab="Age", ylab="dL", type="p", lwd=2, ylim=c(9,10))

abline(h = dl_obs, col="blue")

```

### Get weighted length and optimize alpha1

```{r}
dl_weighted = function(alpha1, weights, tsb = tsbmean, temp = Tmean){
  sum(fishGrowth(alpha1, tsb, temp)*weights)
}

wts = c(0.597721, 0.173708, 0.103396, 0.074332, 0.050843)

dl_weighted(c, weights = wts)

dl_alpha1 = sapply(X = seq(4,8, length.out = 100), FUN = dl_weighted, weights = wts)

plot(dl_alpha1~seq(4,8, length.out = 100), ylab="weighted dl", xlab="alpha1", type="l", lwd=2)
abline(h = dl_obs, col="blue")

```


### Find optimal alpha1

```{r}
error_fun = function(alpha1, weights){
  (dl_weighted(alpha1, weights) - dl_obs)^2
}

optim(par = c(c), fn = error_fun, weights = wts, method="Brent", upper = 8, lower=4)
```

## Growth model as a function of tsb and temperature

### A look at data

```{r}
library(tidyverse)

dat = read.csv("~/Documents/fish_project/growth_model_calib_data/abundance_env_combined.csv")
dat = dat %>% filter(!(is.na(year)))

length_t = mapply(tsb = dat$totb*1e3/1e6, temp = dat$FBT, alpha1 = c, FUN = fishLen)
dl_t = mapply(tsb = dat$totb*1e3/1e6, temp = dat$FBT, alpha1 = c, FUN = fishGrowth)
  
par(mfrow=c(2,2), mar=c(4,4,1,1))
plot(dat$FBT~dat$year, type="l", ylab="Temp (deg C)", xlab="year")
plot(I(dat$totb/1e6)~dat$year, type="l", ylab="Total biomass (MT)", xlab="year")
matplot(y = dat[,2:6], x=dat$year, type="l", lty=1, col=rainbow(5,end=0.75), log="y", ylab = "Abundances", xlab="year")
matplot(y = t(length_t), x=dat$year, type="l", lty=1, col=rainbow(5,end=0.75), log="y", ylab = "Lengths", xlab="year")


```


### Simulated dl as a function of time

```{r}
dl_t_weighted = function(dat, alpha1){
  dl_t = mapply(tsb = dat$totb*1e3/1e6, temp = dat$FBT, alpha1 = alpha1, FUN = fishGrowth)
  wts = dat[,2:6]/dat$Ntot_1_5
  dl_t_wt = rowSums(t(dl_t) * wts)
  dl_t_wt
}

dl_t_wt = dl_t_weighted(dat, c)

par(mfrow = c(3,1), mar=c(4,4,1,1))
plot(dat$FBT~dat$year, type="l", ylab="Temp (deg C)", xlab="year")
plot(I(dat$totb/1e6)~dat$year, type="l", ylab="Total biomass (MT)", xlab="year")
matplot(y = t(dl_t), x=dat$year, type="l", lty=1, col=rainbow(5,end=0.75), log="", ylab = "dL", xlab="year")
points(y = dl_t_wt, x=dat$year, type="l", lty=1, lwd=1, col="black")

```

### Compare mechanistic and statistical dl at default alpha1 

```{r}
tsb_ano = dat$totb*1e3/1e6 - tsbmean
temp_ano = dat$FBT - Tmean
dl_obs = exp(beta1+beta2*tsbmean + beta2*tsb_ano + beta3*temp_ano)

matplot(y = cbind(dl_t_wt, dl_obs), x=dat$year, type="l", lty=1, col=rainbow(2,end=0.75), ylab = "dL", xlab="year")

```

### Optimize alpha1 to maximize match

```{r}
dl_err = function(dat, alpha1){
  dl_t_wt = dl_t_weighted(dat, alpha1)

  tsb_ano = dat$totb*1e3/1e6 - tsbmean
  temp_ano = dat$FBT - Tmean
  dl_obs = exp(beta1+beta2*tsbmean + beta2*tsb_ano + beta3*temp_ano)
  
  sum((dl_t_wt - dl_obs)^2)
}

opt = optim(par = c(c), fn = dl_err, dat = dat, method="Brent", upper = 8, lower=4)
print(opt)

dl_t_wt = dl_t_weighted(dat = dat, alpha1 = opt$par)
tsb_ano = dat$totb*1e3/1e6 - tsbmean
temp_ano = dat$FBT - Tmean
dl_obs = exp(beta1+beta2*tsbmean + beta2*tsb_ano + beta3*temp_ano)

matplot(y = cbind(dl_t_wt, dl_obs), x=dat$year, type="l", lty=1, col=rainbow(2,end=0.75), ylab = "dL", xlab="year")


```