---
title: "Creating and simulating a single fish"
author: "Jaideep Joshi"
date: "28 March 2022"
output: html_document
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = F, echo = F}
library(tidyverse)
library(rfish)
```

### Create a fish with default parameters

A fish can be created by instancing an object of the Rcpp class `Fish`. Once created, the fish **must** be initialized using the `init()` function, which computes parameter values and sets the initial length of the fish. The function `init()` needs 2 arguments, the current total stock biomass (in kilotons) and current temperature (in deg C).


```{r }
fish = new(Fish)
fish$init(tsb = 1.93e3, temp = 5.61)
fish$t_birth
fish$length
```

### Change fish parameters

Suppose we want to create a Fish with custom parameters. 

We can access and set parameters using the `par` object in the fish:

```{r}
fish = new(Fish)
fish$par$c = 6.5

fish$par$c

fish$init(1.93e3, 5.61)
fish$t_birth
fish$length # Notice the change in initial length due to a change in the parameter 'c'
```

Alternativelt, create a parameters object, change the parameters as desired, and assign it to the fish. 

```{r}
# Create parameters object
fish.par = new(FishParams)

# Change specific parameter values
fish.par$s0 = 0.09637

# Assign it to the fish
fish$par = fish.par

fish$par$s0
```


### Get the state of the fish

To get the state of the fish, i.e., its length, age, maturation status, whether it is alive, etc, use the `get_state()` function. The function returns a vector in the following order

+ Year of birth	
+ Age
+ is it mature (T/F)
+ is it alive (T/F)
+ length
+ weight

```{r}
fish = new(Fish)
fish$init(tsb = 1.93e3, temp = 5.61)

fish$get_state()
```

### Simulate the fish

We can now update the status of the fish by calling its demographic functions. 

To update the maturity of the fish, grow it in length over one year, and increment its age by 1 year, we can do
```{r}
fish$updateMaturity()
fish$grow(0, 5.61)  # The arguments are: total stock biomass and temperature
fish$set_age(fish$age+1) # The argument is the new age

fish$length
```


### Simulating a fish over multiple years

Let us now simulate a fish starting at age 1 upto age 30, performing maturation, growth, and age increment every year in that order. We will record the fish length at each timestep so that we can plot the growth trajectory.

```{r}
fish = new(Fish)
fish$init(0, 5.61)
years = 1:31
length = numeric(31)
for (i in years){
  length[i] = fish$length
  fish$updateMaturity()
  fish$grow(0, 5.61)
  fish$set_age(fish$age+1)
}

plot(length~years, xlab="Age", ylab="Length", type="l", lwd=2)
```

Now, since the growth trajectory of a fish depends on the age at which it matures, and since maturation is arandom process, let us simulate lots of fish and look at the possible growth trajectories. 

```{r}
# NEW MODEL
N = 50
plot(x=1, y=NA, xlim=c(0,31), ylim=c(0,250), xlab="Age", ylab = "Length")
for (f in 1:N){
  fish = new(Fish)
  fish$par$s0 = 0.09637
  fish$init(runif(1, min = 0, max = 1.93e3*3), rnorm(1, mean=5.61, sd=3))
  
  dat = data.frame(t_birth=0,	age=0,	isMature=0,	isAlive=0,	length=0,	weight=0)
  dat = rbind(dat, fish$get_state())
  dat = dat[-1,]
  for (i in 1:30){
    dat = rbind(dat, fish$get_state())
    fish$updateMaturity()
    fish$grow(runif(1, min = 0, max = 1.93e3*3), rnorm(1, mean=5.61, sd=2))
    fish$set_age(fish$age+1)
  }
  points(dat$length~dat$age, type="l", col=rainbow(N)[f])
}
```


### Using different models for life processes

Each fish has four life processes:

+ Growth (gro)
+ Maturation (mat)
+ Mortality (mor)
+ Fecundity (fec)

Each of these processes can be simulated either with the model of Dankel et al XXX (old model) or XXX et al XXX (new model). 

To set which model to use, you can set the following four flags to true or false.

```{r}
fish.par$growth_model = 0
fish.par$use_old_model_mat = T
fish.par$use_old_model_mor = T
fish.par$use_old_model_fec = T
```

Here is the growth trajectory of the fish using the model in Dankel et al XXX.

```{r}
fish = new(Fish)
fish$par$growth_model = 0
fish$init(1.93e3, 5.61)


dat = data.frame(t_birth=0,	age=0,	isMature=0,	isAlive=0,	length=0,	weight=0)
dat = rbind(dat, fish$get_state())
dat = dat[-1,]
for (i in 1:30){
  dat = rbind(dat, fish$get_state())
  fish$updateMaturity()
  fish$grow(1.93e3, 5.61)
  fish$set_age(fish$age+1)
}
plot(dat$length~dat$age, type="l", xlab="Age", ylab="Length", col="black")
```


### Compare growth and mortality based on old and new models

```{r}

plot(x=1, y=NA, xlim=c(0,31), ylim=c(0,200), xlab="age", ylab = "length")

# NEW MODEL
mat_ages = numeric(0)
mort = data.frame(a=0, m=0, l=0)
N = 100
for (f in 1:N){
  #cat("f = ", f, "\n")
  fish.par = new(FishParams)
  fish.par$s0 = 0.09637
  # fish.par$Bhalf_growth = 1e11
  
  fish = new(Fish)
  fish.par$flag = 1
  fish.par$growth_model = 1
  fish$par = fish.par
  fish$init(1.93e3, 5.61)
      
  dat = data.frame(t_birth=0,	age=0,	isMature=0,	isAlive=0,	length=0,	weight=0)
  dat = rbind(dat, fish$get_state())
  dat = dat[-1,]
  for (i in 1:30){
    #cat("rmat prob: age = ", fish$age, ", p = ", 1 / (1 + exp( -(fish$length - (1.604122 * fish$age + 18.399575)) / 1.93031)), "\n")
    #cat("f = ", f, "\n")
    dat = rbind(dat, fish$get_state())
    mort = rbind(mort, c(a=fish$age,  fish$naturalMortalityRate(), fish$length))
    fish$updateMaturity()
    fish$grow(1.93e3, 5.61)
    fish$set_age(fish$age+1)
  }
  mat_ages = c(mat_ages, which(dat$isMature>0)[1])
  points(dat$length~dat$age, type="l", col=rainbow(N)[f])
}
table(mat_ages)

# OLD MODEL
N=1
for (f in 1:N){
  #cat("f = ", f, "\n")
  fish.par = new(FishParams)
  fish.par$s0 = 0.09637
  # fish.par$Bhalf_growth = 1e11

  fish = new(Fish)
  fish.par$flag = 1
  fish.par$growth_model = 0
  fish$par = fish.par
  fish$init(1.93e3, 5.61)
  
  dat = data.frame(t_birth=0,	age=0,	isMature=0,	isAlive=0,	length=0,	weight=0)
  dat = rbind(dat, fish$get_state())
  dat = dat[-1,]
  for (i in 1:30){
    #cat("rmat prob: age = ", fish$age, ", p = ", 1 / (1 + exp( -(fish$length - (1.604122 * fish$age + 18.399575)) / 1.93031)), "\n")
    #cat("f = ", f, "\n")
    dat = rbind(dat, fish$get_state())
    mort = rbind(mort, c(a=fish$age,  fish$naturalMortalityRate(), fish$length))
    fish$updateMaturity()
    fish$grow(1.93e3, 5.61)
    fish$set_age(fish$age+1)
  }
  mat_ages = c(mat_ages, which(dat$isMature>0)[1])
  points(dat$length~dat$age, type="l", lwd=2, col="black")
}
```

### Maturation



### Mortality

```{r}
source("../tests/ref/parameters.cod.R")
M<-c(M,rep(M[length(M)],times=ages-length(M))); names(M)<-1:ages
with(mort %>% filter(a<=20 & a>1), plot(m~a, ylim=c(0,1), xlab="Age", ylab="Mortality rate"))
points(M, type="l", col="blue")

length_age = mort %>% group_by(a) %>% summarize(l = mean(l))

```

